### 2018-06-12 paper/paper.tex
reorganised for efficiency

\documentclass[11t, a4paper, twocolumn]{article} 
\input{structure.tex}
\addbibresource{references.bib}
\title{Estimating real-time highstreet footfall from Wi-Fi probe requests}
\author{
	\authorstyle{
		Balamurugan Soundararaj\textsuperscript{1}, 
		James Cheshire\textsuperscript{1} and 
		Paul Longley\textsuperscript{1}}
	\newline\newline
	\textsuperscript{1} \institution{
		Department of Geography, 
		University College London, 
		United Kingdom } }
\date{\today}
\begin{document}
	\maketitle
	\thispagestyle{firstpage}
	\textbf{Abstract -}
		\input{sections/abstract}
	\section{Introduction}\label{introduction}
		\input{sections/introduction}
	\section{Previous Work}\label{previous_work}
		\input{sections/previous_work}
	\section{Methodology}\label{methodology}
		\input{sections/methodology}
	\section{Pilot Study}\label{pilot_study}
		\input{sections/pilot_study}
	\section{Main Study}\label{main_study}
		\input{sections/main_study}
	\section{Discussion}\label{discussion}
		\input{sections/discussion}
	\section{Conclusion}\label{conclusion}
		\input{sections/discussion}
	\printbibliography[title={References}]
\end{document}