### 2018-06-24 paper/sections/abstract.tex

The accurate measurement human activity with high spatial and temporal granularity is crucial for us to understand the structure and function of built enviroment.
With increasing mobile ownerships, the Wi-Fi probe requests generated by these devices can be an excellent source of data for such measursement.
The major challanges in using Wi-Fi technology for such purpose are precisely delineating the field of measurement and arriving at accurate estimates without compromising the privacy of the users of the mobile devices.
In this paper we demonstrate that, with the application of class intervals and a novel graph based technique, we can overcome the challanges and reliably measure real-time footfall at retail highstreets in London.