### 2018-06-29 paper/sections/abstract.tex
Paul's comments included

The accurate measurement human activity with high spatial and temporal granularity is crucial for understanding the structure and function of built environment.
With increasing mobile ownership, the Wi-Fi `probe requests' generated by mobile devices can act as an cheap, scalable and real-time source of data for such establishing such measures.
One of the major challanges in using Wi-Fi technology for estimating human activity is eliminating biases caused by an uncertain field of measurement and changing technological landscape without compromising the privacy of the users of the mobile devices.
In this paper we demonstrate that, with the application of class intervals and a novel graph based technique, we can overcome the challanges and reliably measure real-time pedestrian footfall at retail highstreets.