### 2018-06-29 paper/sections/introduction.tex

New and developing technologies today provide the infrastructure over which movements and interactions can be measured and monitored in the `sentient city' \citep{amin2017seeing}.
These include mobile phone networks, which can triangulate user locations relative to networks of masts, use of GPS to locate users of social media services, and WiFi beacon connectivity to access the Internet.
These technologies offer differing levels of spatial precision, where mobile telephony and WiFi generally being less reliable and offer lower precision than GPS to the end users, while being more advantageous for broader mobility studies \citep{pinelli2015comparing}.

There has been considerable research into the usefulness of these technologies to understand patterns of movement in cities in near real time \citep{candia2008uncovering, gonzalez2008understanding, calabrese2013understanding}. Most of this research has focused upon technical specification of accuracy or precision \citep{song2010limits, lane2010survey}, with somewhat less attention devoted to the ways that the characteristics of the technologies and of their human users conspire to create possible bias in the inclusiveness of usage patterns.
This is perhaps most manifest in the plethora of papers that analyse the distribution and content of publicly available Twitter social media feeds, which likely constitute an extremely biased and self-selecting subset of users that geo-enable their posts, drawn from a user base that itself bears no identified correspondence to any known background population \citep{lansley2016deriving}. 
Even analysis of mobile phone data, usually derived from industry players that have significant market share and user bases that may be representative of local populations, may exclude groups such as tourists from distand origins or subscribers to third party services that share distinctive characteristics \citep{di2016mind}
These examples illustrate the issues that underpin the assembly and analysis of consumer data, which can be considered as a distinctive class of Big Data that arise from the interactions between humans and customer-facing organisations such as retailers, domestic energy suppliers, transport providers and suppliers of social media and communications \citep{cdrc2018consumer}.
Consumer data account for an ever-increasing real share of all of the data that are collected of citizens, but a fundamental characteristic of consumer-led markets is that no single provider has a monopoly in market provision, and therefore issues of market share and segment generate bias in analysis.
The source and operation of this bias is unknown in the absence of extensive and context sensitive attempts to triangulate consumer data with data of known provenance relating to clearly defined populations \citep{lansley2016deriving}.
In similar ways to other classes of Big Data, consumer data are best throught of as digital ‘exhaust’, or a by-product created by or harvested from consumer transactions.

In this paper, we evaluate the value of data collected from a Great Britain wide network of 800 devices, installed in order to characterize the footfall patterns of a scientifically balanced sample of retail centres.
These devices are located in shop windows and record the probes emitted by mobile phones and other WiFi enabled devices.
These are deemed to be consumer data because devices carried by consumers routinely probe for a consumer service, specifically a WiFi connection.
Monitoring the probes from such devices provides an indication of the presence of their users, regardless of whether or not Internet connectivity is established.
Our core motivation is to appraise the usefulnessof WiFi probe records harvested from our network of sensors in order to indicate levels of pedestrian activity.
More broadly still, in our future research we intend to classify the nationwide network of footfall profiles as part of a programme of research to understand the form and functioning of retail areas at a time of far-reaching structural change for the retail industry.

To this end, it is important to first undertake a thorough conceptual and technical appraisal of our consumer data source.
In technical terms, screening of information present in the `probes' and classifing them based on their characteristics is essential in order to remove the ones emitted by devices that do not indicate pedestrian activity, such as network enabled printers and other fixed devices.
Related to this, a method to figerprint Wi-Fi probes is necessary to remove probes from individuals’ devices that in conceptual terms should not be considered part of footfall – as when, for example, an employees is seated in an office within range of the sensor device.
A calibration of sensor measurement is also essential on two grounds: first, individuals may carry multiple devices, or no device at all; and second, the positioning and orientation of the sensor in the retail unit may lead to systematic over- or under-enumeration.
These sources of bias in measurement must be accommodated by manual recording of footfall at each location and generalization of these sample survey results to all locations and time periods.
As we describe in detail below, manual validation of the data needs to be undertaken in parallel with technical profiling of the mix of consumer devices that probe our sensors, since the effectiveness of data cleaning procedures discussed in this paper differ between individual locations and configurations.