### 2018-06-30 paper/sections/abstract.tex
start proofreading

The accurate measurement of human activity with high spatial and temporal granularity is crucial for understanding the structure and function of the built environment.
With increasing mobile ownership, the Wi-Fi `probe requests' generated by mobile devices can act as a cheap, scalable and real-time source of data for establishing such measures.
The two major challanges we face in using these probe requests for estimating human activity are, filtering the noise generated by the uncertain field of measurement and clustering anonymised probe requests generated by the same devices together without compromising the privacy of the users.
In this paper we demonstrate that we can overcome these challanges by using class intervals and a novel graph based technique for filtering and clustering the probe requests which in turn, enables us to reliably measure real-time pedestrian footfall at retail highstreets.