### 2018-11-20 paper/sections/conclusion.tex

Sentient technologies make possible measurement of the human activities that are
the life blood of the smart city. Yet the data that they harvest are frequently
relevant only to sub-groups within society that avail themselves of particular
goods and services – such as social media applications, transport modes or
retail offers. In each of these cases, it is necessary to remember that the
resulting data are a by-products of consumer transactions, and will as a
consequence, only pertain to users of the relevant good or service. If the smart
city is to be socially inclusive, it therefore follows that sentient data must
represent entire populations, whether by design or by triangulation with
external, population wide, sources. This usually poses a non-trivial task, since
the ebbs and flows of smart device-enabled citizens rarely pertain to any
clearly defined population in either administrative or functional terms Massam
1973.

Our objective here has been to collect, rather than re-use, data on smart city
functioning, by recording Wi-Fi probes and ultimately reconciling them with
manual counts in order to infer ambient populations. The internal validation
methodology set out in the technical sections of this paper, allied to external
validation from pedestrian counts, renders the method inclusive and robust when
recording activity levels in retail centres in real time. We have described the
collection and processing of a novel consumer Big Dataset that enables valid
measures of levels of footfall activity which has been scaled across a Great
Britain wide network of sensors CDRC 2016.  In both conceptual and technical
terms, it illustrates the ways in which passively collected consumer data can be
‘hardened’ to render them robust and reliable by using related procedures of
internal and external validation.

Internal validation addresses the issues of screening out device probes that do
not indicate footfall, and the further screening of device probes to
‘fingerprint’ the effects of MAC randomization. It is important to note that the
filtering process work based solely on the information present in the probe
requests and their temporal distribution. This ensures that although the mobile
devices were uniquely identified, there was no further personal data generated
by linking the probe requests to the users of the mobile devices. This method
essentially gave us a way to estimate the footfall in real-time without
identifying or tracking the mobile devices themselves. External validation then
entails reconciling adjusted counts with the footfall observed at sample
locations. This procedure makes it possible to generalise from locations at
which manual footfall surveys are conducted to all others in the system, and to
develop a classification of device locations that are more or less susceptible
to noise generation.

This Wi-Fi based footfall counting methodology offers a large number of
applications and benefits for real time spatial analysis. Since Wi-Fi based
sensors are inexpensive and the data model is scalable, it is possible to use
this methodology for a large network of sensors to gather granular data on
pedestrian footfall. A snapshot showing week’s worth of precise footfall in area
around Charring cross, London is shown in Figure 7 to demonstrate the potential
for such dataset. Projects such as SmartStreetSensors CDRC 2016, may utilise
this methodology to overcome the challenges introduced by the implementation of
MAC address randomisation.

The vicissitudes of MAC randomisation, and the provisions of privacy legislation
such as EU General Data Protection Regulations mitigate against tracking
individuals across the smart city using this approach.  This can be modelled
using agent-based methods Heppenstall et al 2012, however. In our own research
we have also begun to link store time-lagged till receipts to footfall, and have
used such data to better understand the dwell times that characterise such
different retail uses as stores with window displays and fast food restaurants.
Such analysis not only provides a more nuanced picture of movement through
retail areas, but also enables valorisation of micro sites within retail
centres. In the UK, for example, this is of immediate practical importance in
evaluating business rates on properties, and has still wider implications for
the setting of retail unit rental values. There are obvious extensions to
understanding the ebbs and flows of activities in the 24 hour smart city.

More broadly still, extensions to this strand of smart city research are likely
to seek to differentiate the quality of different elements within footfall
according to mission e.g. travel to adjacent workplace zones, leisure, etc and
personal characteristics such as spending power. In this respect, future
research may not only simulate linkage of harmonised footfall counts between
sensor locations, but also link these in turn to disaggregate origin-destination
matrices for bikeshare and other public transport modes.  Our own investigations
will consider these and other challenges to understanding the functioning of the
sentient city.