### 2018-11-20 paper/sections/pilot_study.tex

\begin{table}
\tbl{Comparison of clustering algorithms with a sample of 40000 probe requests}
{\begin{tabular}{lcc} 
\toprule
	 Algorithm					& Time (s) 	& MAPE (\%) \\
\midrule
	 Quantile					& 0.002 	&  27\% \\
	 K-Means			 		& 0.007 	& -23\% \\
	 Hierarchical Clustering	& 172.520 	& -9\% \\
	 Bagged Clustering 			& 0.135 	& -30\% \\
	 Fisher 					& 3.034 	& -30\% \\
	 Jenks Natural Break 		& 556.279 	& -30\% \\
 \bottomrule
\end{tabular}}
\label{classification-table}
\end{table}

To start we designed a small pilot study to validate the filtering and
clustering methodology against the scale and complexity of data collected on an
open public area such as a retail high street.  We also aim to find the
algorithm which is best suited for classification of signal strengths as 'low'
and 'high' to filter out the background noise.  The data was collected at Oxford
Street, London on 20 December 2017 from 12:30 to 13:00 hrs, where Wi-Fi probe
requests were collected using the sensor described in Chapter \ref{methodology}
and pedestrian footfall was manually recorded using the Android app
\citep{bala2018clicker}.  Being located at one of the busiest retail locations
in the United Kingdom, the Wi-Fi sensor captured approximately 60,000 probe
requests during the half hour period, and 3,722 people were recorded manually
walking on the sidewalk during that time.

As a first step we just aggregated the probe requests by their MAC address for
every minute to generate a minute by minute count of the number of people around
the sensor assuming each MAC address corresponds to a mobile device and hence a
pedestrian.  We then compare this preliminary `footfall' count to the actual
number of pedestrians recorded manually to check for the robustness. We use Mean
Absolute Percentage Error (MAPE) measure of robustness of the count since it
provides a simple and quick measurement while the street conditions ensure that
there are no intervals without any footfall.  We find that the MAPE in the raw
counts compared to the ground truth is around 425\%.  This suggests the presence
of large amount of noise in the data which might be generated due to sources of
uncertainties discussed in Chapter \ref{methodology} thus demonstrating the need
for filtering the data.

We then classified the probe requests as ones with "high signal strength" and
"low signal strength" using various one dimensional clustering algorithms such
as k-means, quantile, hierarchical clustering, bagged clustering, fisher and
jenks natural breaks.  The results are shown in Table
\ref{classification-table}.  We found that while hierarchical clustering and
jenks gave us fairly low errors, they were too resource intensive for practical
use with a larger dataset.  We also found that k-means gave quickest results
with the lowest MAPE closely followed by quantile algorithm.  The cut-off point
or threshold for the collected data with which we could classify them as high
and low was -71 dBm.  We then removed all the probe requests which reported `low
signal strength' and repeated the same aggregation process as before to produce
footfall count.  This process resulted in a footfall count with a net MAPE of
30\%.  Though the results are encouraging we are still not completely confident
that our filtering process is indeed removing noise or has any correlation the
configuration of sensor or position of the mobile devices.  These need to be
addressed with a larger survey with multiple location of varying orientations.

\begin{figure}
\begin{center}
\vspace{10pt}
\includegraphics [width=\linewidth,trim=10 30 10 75,clip]
	{images/pilot_clustering_params.png}
\caption{The clustering process was repeated with both increasing sequence
	number threshold ($\alpha$) and time threshold ($\beta$), until we arrived
	at the lowest parameters where the know device (black line) is clustered as
	a single device.}
\label{pilot_clustering_params}
\end{center}
\end{figure}

\begin{figure}
\centering
\subfigure[Clustering probe requests based on increasing sequence numbers 
	present in them.]
	{\resizebox*{0.45\linewidth}{!}
	{\includegraphics{images/pilot_clustering.jpeg}}}
\hspace{20pt}
\subfigure[Comparison of sensor based counts with the ground truth.]
	{\resizebox*{0.45\linewidth}{!}
	{\includegraphics{images/pilot_comparison.jpeg}}}
\caption{The results of the pilot study demonstrating the validity of the
	methodology.} 
\label{pilot_results}
\end{figure}

The next challenge was to identify probe requests which are generated by the
same device irrespective of MAC randomisation process.  We use the algorithm
defined in Chapter \ref{methodology} and assign a unique identifier or signature
to each probe request independent of the MAC address.  Since there is neither
prior research nor documentation on a universal behaviour of phones in
randomising their MAC addresses, we use the stationary device - the one used for
manual counting as a reference and find out the suitable time threshold,
$\alpha$ and threshold for sequence numbers, $\beta$ to be 16 seconds and 60
respectively via trial and error. This process is shown in Figure
\ref{pilot_clustering_params} This process is done on top the filtering done
based on signal strength and only for the probe requests with randomised MAC
addresses.  Figure \ref{pilot_results} shows the results of this clustering
process on a small set of randomised probe requests.  The probe request with
different randomised MAC address is shown by the colored points and the line
joining them shows the ones belonging to the same cluster hence expected to be
generated by the same device.  We finally aggregate the probe requests as before
but with the device signature rather than just MAC addresses this results in a
footfall count with a MAPE of -18\%.  A comparison of minute by minute counts
resulting from different filtering processes along with the ground truth is
shown in Figure \ref{pilot_results} showing the promising effectiveness of the
methods.

To conclude, from the pilot study we found that both filtering and clustering
methods we devices work on complex real world data and results in a final
pedestrian counts within a MAPE of 20\%. We also found `k-means' and `quantile'
are best algorithms for classifying signal strengths and the threshold for time
and sequence numbers for the clustering algorithm is around 16 and 60
respectively.