### 2018-11-20 paper/sections/previous_work.tex

In the past decade Wi-Fi has emerged as one of the most commonly used technology
in providing high speed internet access to mobile devices such as smartphones,
tablets and laptops in public and private spaces. This has resulted in multiple
Wi-Fi networks being available at almost every location in dense urban
environments. Traversing through this overlapping mesh of Wi-Fi networks, modern
mobile devices with Wi-Fi antennae regularly broadcast a special type of signal
known as `Probe Requests' in order to discover Wi-Fi networks available to them.
This helps these devices to connect and switch between the Wi-Fi networks
seamlessly.

Probe requests are low level signals standardised by IEEE 802.1b/g specification
\citep{ieee2013} as the first step in establishing a Wi-Fi based connection
between two devices and is implemented in any Wi-Fi capable device irrespective
of the manufacturer or the model. This ubiquity and standardisation makes them
an excellent source of open, passive, continuous, and wireless data generated by
Wi-Fi capable devices present at any given time and location. Considering the
unprecendented levels of mobile device ownership in recent years, we can, in
turn use this data to understand the population distribution in highly dynamic
urban environments with high spatial and temporal granularity
\citep{freud2015,konto2017}. While a Wi-Fi based method to collect data offers
us various advantages such as, easy scalability and efficiency in terms of cost
and time, It also introduces few systematic biases, uncertainities in the
collected data along with the serious risk of infringing on the privacy of the
mobile users. In this paper, using a set of probe requests and manual counts
collected at various high street locations across London, we demonstrate that
pedestrian footfall at these locations can be estimated with considerable
precision and accuracy while protecting the privacy of the pedestrians.

Though, unlike GPS, the location of the Wi-Fi enabled mobile device cannot be
directly inferred from Wi-Fi, there are reliable methods to triangulate the
location of mobile devices from the locations of known access points(AP) and the
signal strength reported by them
\citep{he2003range,moore2004robust,lamarca2005place}. This can overcome the
usual shortcoming of GPS, which struggles for precision and accuracy in indoor
and densely built environments
\citep{zarim2006,kawaguchi2009wifi,xi2010locating}. Utilising this, we can
easily and quickly estimate trajectories of the mobile devices
\citep{Sorensenlocation,musa2012tracking} which can be used similar to the GPS
trajectories to understand individual travel patterns \citep{reki2007,Sap2015},
crowd behaviour \citep{abedi2013bluetooth,mowafi2013tracking}, vehicular
\citep{lu2010vehicle} and pedestrian movement
\citep{xu2013pedestrian,fukuzaki2014pedestrian,wang2016gait}. Such data can
also be used in transportation planning and management to estimate travel time
\citep{musa2011wiflow} and real time traffic monitoring
\citep{abbott2013empirical}. Using techniques demonstrated by
\cite{franklin2006passive} and \cite{pang2007802} along with information present
in the probe requests one even model interactions between the users
\citep{cheng2012inferring,barbera2013signals,cunche2014know,cunche2014linking}
such as predicting which of them are most likely to meet again
\citep{cunche2012know}. Using the semantic information present in these probe
requests it even is possible to understand the nature of population at a large
scale \citep{di2016mind}.

Though extensive research has been carried out on this subject with feasible and
favorable results, in recent years, one of the major challenges faced in such
attempts has been the increasing attempt by mobile phone manufacturers to
protect their users’ privacy by anonymising the globally identifiable portion of
the probe requests, \citep{green2008}. Various methods have been devised to
overcome this anonymisation process such as estimating the device model
information from a known dataset of manufacturers and device behaviours
\citep{martin2016decomposition}; Scrambler attack using a small part of the
physical layer specification for Wi-Fi \citep{vo2016,bloessl2015scrambler}; and
timing attack where the packet sequence information along with information
elements present in the probe request frame is used
\citep{matte2016,cheng2016can}. A combination of these methodologies has been
proven to produce de-anonymised globally unique device information
\citep{vanhoef2016,martin2017}. These approaches usually result in serious risk
of breach of privacy of the users of the mobile devices by revealing their
identifiable personal information.

There is a clear gap in research for exploring methodologies for estimating the
number of unique mobile devices from a set of anonymised probe requests, without
the need to reveal their original device information. Such technique has
various applications such as uncovering the urban wireless landscape
\citep{rose2010mapping}, revealing human activity at large scales
\citep{qin2013discovering}, estimating pedestrian numbers in crowds
\citep{schauer2014estimating,fukuzaki2015statistical} and even counting people
in hyper local scales such as queues \citep{wang2013measuring} With enough
infrastructure to collect such information we can even aim to generate a
real-time census of the city \citep{konto2017}. With this background we set out
to device and implement a methodology to reliably estimate human activity such
as pedestrian footfall from Wi-Fi probe requests without risking the breach of
privacy of the users involved.