### 2018-11-25 review/response.md
Sun 25 Nov 22:25:41 GMT 2018

First review comments
=====================

Title : Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests

Manuscript id : IJGIS-2018-0340

Comments dated : 28-Oct-2018

1 Editor
--------

1.1 Overall Comments
---------------------

1. While this is an important and timing topic dealing with data uncertainty
   and privacy in the mobile age, the reviewers have raised concerns on
	- Clarity of the methodology
	- Lack of discussion of wifi-based human activity studies in the big data
	age
	- Proper use of terminology
	- Performance of the algorithm etc.

2. In lieu of these comments, they have recommended to make substantial
revisions to your manuscript

3. Also, since this is a "smart cities" related special issue, I ask you to
	- Add more discussions on smart cities in the introduction and conclusion
	sections and
	- Link your research more tightly to this theme.

> Thank you for the very helpful comments of the reviewers of our paper/ As
> requested, we have responded to all comments without adding significant
> length to the paper. We have also extended our discussion of how effective
> measurement of footfall is relevant to understanding the functioning of the
> Smart City. If any of what follows and what has been uploaded is unclear,
> please do not hesitate to contact me.
> 
> Yours sincerely,  
> Balamurugan Soundararaj,  
> (on behalf also of Paul Longley and James Cheshire)

1.2 General Instructions
------------------------

1. Please try to avoid making the paper longer (keep in mind that the average
   length of papers in this journal is 18-20 published pages). It is important
that you ensure that the spelling and grammar is of a high standard.

2. Please also highlight the changes to your manuscript within the document by
   using the track changes mode in MS Word or by using bold or colored text.

3. To revise your manuscript, please select the link below,
   https://mc.manuscriptcentral.com/ijgis?URL_MASK=a32caf3bbd49412cac8d1cfa9d7ebaf8
   which will take you your decision letter, and where you need to resubmit
   your manuscript.

4. If you have used the on-line LaTeX template, upload all those files and a
   pdf.

5. Please check your English spelling and grammar.

6. There are two hard limits that you need to bear in mind, 
	- Image size - Please be aware that the total number of pixels (height X
			width) in each image must be less than 40,000,000 (40 mp).
	- Total file size - please make sure that the combined, uncompressed, file
	sizes do not exceed 20Mb. The internal LZW compression, which is an option
	in TIF, is acceptable however.
	- Please do not save B/W or Greyscale images as RGB.

7. Equations and tables need to be in an editable format.

8. You need to check that the paper title in ScholarOne is the same as the title
   on the paper and that the Title page contains all authors’ names and
   affiliations as you would want them to appear.

9. When submitting your revised manuscript, you will be able to respond to the
   comments made by the reviewers in the space provided. You can use this space
   to document any changes you make to the original manuscript. In order to
   expedite the processing of the revised manuscript, please be as specific as
   possible in your response to the reviewers.

10. Please delete any redundant files before completing the submission.

> Thank you for the detailed instructions. All of these have been taken in
> consideration and checked for creating the revised manuscript for the final
> submission.

2 Reviewer 1
------------

2.1 Overall Summary
-------------------

1. This paper is about the estimation of pedestrian footfall using Wi-Fi signal
   emitted by mobile devices. It deals with the problem of estimating the
   footfall size based on a noisy and biased data. More specifically it
   addresses the following issues that may lead to over counting:

 - MAC randomization that may cause a device to generate several MAC addresses;
 - Signals received from outside the interest zone; unequal adoption of Wi-Fi
 smartphone in the population. To address this issue, a number of techniques
 are introduced:
 - clustering received signal strength to remove device that are outside the
 monitored area;
 - clustering probes based on sequence number and timing;
 - finally an adjustment factor is applied to correct the estimated value.

2. The adjustment factor is calibrated for each location using a manual
   counting. The author conducted a pilot experiment to identify the clustering
   algorithms and parameters, and then they conducted a second range of
   experiments at different location over an extended period of time. During
   each of those experiments, ground-truth count was collected manually. The
   results show that the presented approach can significantly reduce the
   counting error compared to a naive solution consisting in counting the
   distinct number of MAC observed. Nevertheless, there are cases (2) where the
   error is still large (50%), even if in other cases (1) the error is only of
   9%.

> This is a correct surmisal of the remit of the paper. We have also added
> material to the introduction and conclusion of the paper to better link this
> remit to the Smart Cities discourse.

2.2 General Comments
--------------------

1. The idea of clustering frames using sequence number has been introduced in
   [1]. This work needs to be cited when introducing the proposed clustering
   algorithm.

   [1] Mathy Vanhoef, Célestin Matte, Mathieu Cunche, Leonardo Cardoso, Frank
   Piessens. Why MAC Address Randomization is not Enough: An Analysis of Wi-Fi
   Network Discovery Mechanisms. ACM AsiaCCS, May 2016, Xi’an, China. 2016,

> Thank you. We have included a citiation to the source, as well as one other
> recent addition to this fast-developing literature.

2. The author considered that the hashing of MAC address is a good way to
   protect users privacy, as some MAC address collected by the device may not be
   random. However hashing of MAC address cannot be considered as a proper
   anonymization technique (see [2] and
   http://webpolicy.org/2014/03/19/questionable-crypto-in-retail-analytics/
   ); at best it can be considered as an obfuscation. [2] Levent Demir, Mathieu
   Cunche, Cédric Lauradoux. Analysing the privacy policies of Wi-Fi trackers.
   Workshop on Physical Analytics, Jun 2014, Bretton Woods, United States. ACM,
   2014

> Thank you for pointing this out. In the revised version of the paper we have
> made reference to this discussion, while also making clear the anonymisation
> is ensured by the system architecture that processess and curates the data.
> Specifically we rotated the random salt used to hash the MAC addresses
> regulary to prevent any kind of tracking of the users across a longer period
> of time with which the hashing process is carried out.

3. There is a number of commercial solutions doing counting based on Wi-Fi
   signal. Although the detail of their solution is not public, it would be
   worthwhile to acknowledge their existence.

> Thank you - this has been done.

4. In section 3.2, it is explained that data is divided into two sets
   corresponding on random and non-random MAC. Which method is used to perform
   this classification ?

> The identification of a MAC address as randomised or non-randomised is done
> looking at the second character of the vendor part of the MAC address. The
> addresses with E, A, 2, 6 at these locations are identified to be randomised
> MAC addresses. This has been included in the revised paper.

5. How was the sensor positioned compared to the area monitored used for ground
   truth ? A more detailed description is required regarding the positioning of
   the sensor and the size and topology of the surveyed area.

> This is a good point, and is very important for our validation work. We have
> included a more detailed and explicit discussion of the way that field
> observation was used to calibrate and cross validate the Wi-Fi results.

2.3 Clarifications & Corrections
--------------------------------

1. The topic addressed in this work is interesting and timely. The proposed
   approach is sound and shows promising results. The fact that the authors
   conducted field experiments to evaluate their solution is a positive point.
   Yet before publication, I believe that there are a number of points that
   needs to be clarified or corrected. Regarding the technical aspects I have a
   number of minor remarks:

> Thank you very much for these very positive comments. The paper reports on
> extensive methodological research and field implementation and is, we believe,
> cutting edge in its finding. It also has important implications for our
> understanding of the functioning of smart cities, and these implications are
> discussed in greater detail in the revised version.

2. "WiFi beacon technologies to access the Internet". In Wi-Fi, beacons are
   management frames used by Access Point for service discovery. They do not
   provide Internet access. "Wi-Fi Access Point to access ..."

> Thank you. We have corrected and clarified the point.

3. "Wi-Fi antennae regularly broadcast a special type of signal" -> the antenna
   is just the part ensuring that the signal is correctly emitted, but it is not
   the source. In this context it is better to use a more general concept and
   refer to the "Wi-Fi network interface" (or Wi-Fi interface) as the the
   element broadcasting a signal.

> Thank you. Point clarified in the revised version.

4. The IEEE 802.11 b/g is just a subpart of the 802.11 specification. Here you
   can refer to the 802.11 specifications as a whole : [“802.11-2012 - IEEE
   Standard for Information technology–Telecommunications and information
   exchange between systems Local and metropolitan area networks–Specific
   requirements Part 11: Wireless LAN Medium Access Control (MAC) and Physical
   Layer (PHY) Specifications,” IEEE Std 802.11-2012 (Revision of IEEE Std
   802.11-2007), pp. 1–2793, Mar. 2012.

> Thank you. Noted and added to the revised version.

5. "the first step in establishing a Wi-Fi based connection " is also known as
   service discovery

> Thank you. Added to the revised version

6. "Wi-Fi functionality has been turned off by the user", regarding this point
   you may be interested in :

7. "Media Access Control (MAC) address which is an unique identifier for the
   wireless hardware of the mobile device"  The transmitter field of the 802.11
   frame can indeed contain a unique MAC address  but can also contain a random
   one as noted by the author. Here it would be better to avoid stating that the
   content of this field is always a unique MAC.

> Thank you. Again, this has been noted in the revised version of the paper.

8. "The strength of the signal which transmitted the request.". In fact the
   Received Signal Strength Indicator, a value providing an indication on the
   strength of the received signal as seen by the receiver, not the strength at
   the emitter side.

> Thank you. This distinction has been noted in the revised version of the
> paper.

9. Could you provide more information on the application used for counting
   (Clicker)

> This simple application was developed specifically for the project for Android
> platform. More details on the application has been included in the revised
> manuscript.

10. "Though the recycling of sequence number r after 4000 leads to multiple
	classifications reported on single device, the magnitude of error is greatly
	reduced." This sentence is not clear. And the cycle length is not 4000 but
	4096 = 2^12.

> Thank you. This point has been boted and accommodated into the revised
> version.

11. Regarding the issue of cycling sequence number, one could use a distance
	with a modulo : di,j = di - dj mod 4096.

> Thank you very much for the suggestion. From tests carried out on sample of
> 1910 probe requests from google phones in the pilot study we found that around
> 0.5% reset their mac addresses. Which in our opinion doesn't influence the
> results in any significant way considering the context.

2.4 Typos and other remarks
--------------------------------

1. Challanges

2. "These are deemed to be consumer data because devices carried by consumers
   routinely probe for a consumer service, specifically a Wi-Fi connection.
   Monitoring the probes from such devices provides an indication" This sentence
   is not clear.

3. "Though Wi-Fi F is a ‘location-less’ technology, ", please clarify what you
   mean by that.

4. "privacy infringement" may not be the best word as infrigement relate to a
   law or a regulation. Privacy breach or leak may be a better choice.

5. Wireshark and not WireShark.

6. "We hypothesize that an adjustment factor could be arrived at for each
   location of data collection, ""This calibration can be carried over
   periodically and the frequency to improve the quality of the estimation."
   something wrong in both sentences.

> Thank you for picking up these minor lapses. All have been corrected or
> clarified in the revised version of the paper.

3 Reviewer 2
------------

3.1 Overall Summary
-------------------

1. This paper addressed the data quality issue of Wi-Fi-based probe requests in
   estimating pedestrian counts. The authors proposed several methods to improve
   the accuracy of estimating the number of unique mobile devices from a set of
   anonymized probe requests without revealing their original device
   information.

> This is correct, although the intended emphasis upon disclosure is a feature
> rather than a principal motivation. We have also added further discussion of
> how these principally technical issues relate to smart city data
> infrastructure issues.

2. This is a crucial topic in mobile data uncertainty, and the manuscript has
   the potential of becoming a useful reference in the field. However, it was
   not organized effectively and missing many details in the methodology. Before
   addressing these problems, it is difficult to assess the reliability and
   scientific merit of the case study.

> Thank you very much for this endorsement of the motivation for our paper.  We
> hope that following responses and the amendments to the paper will allay your
> concerns about the methodology and application.

3.2 General Comments
--------------------

1. First, the literature review section adequately covered Wi-Fi-based human
   activity studies, but it is in lack of an overview of modeling human activity
   in the big data era. I would suggest the authors add an overarching paragraph
   (or a sub-section) in Section 2 discussing the importance and challenges of
   modeling human activities and urban dynamics based on various types of big
   geodata, such as mobile phone records, location-based social media, Bluetooth
   data, etc.

> This is a welcome comment, and we have added material in the introduction and
> conclusion as to the value of new Big Data sources for developing
> activity-based representations of behaviour and circumstances in the smart
> city, This chimes with much of our related research, but we write cognisant of
> the article length restrictions of IJGIS.

2. The conclusion did not address most of the important results from the case
   study and should be expanded.

> Yes, we agree, and some wider conclusions about the implications of our
> research for better understanding the functioning of the smart city have now
> been added.

3.3 Corrections and Clarifications
----------------------------------

Second, the methodology should be better clarified. There are many details
missing or in need of a better justification. For example,

1. P4 l50, without providing any explanation, the authors hypothesized that
   sequence number and length of the packet are sufficient to estimate the
   number of unique devices. This argument was not elaborated until 2.5 pages
   later in section 3.2.

> This material has now been amended and resequenced.

2. P6 l27, please provide more information regarding how you defined the
   threshold for low/high signal strength when eliminating the background noise.
   This was also unclear in the case study on P8, where the authors repeatedly
   mixed up classification and clustering. Neither k-means nor hierarchical
   clustering is considered a “classification algorithm” as claimed in p8 l26.
   Very few details were provided regarding the parameter setting of these
   analyses.

> Thank you. The reference to clustering has been amended. The specifics of the
> thresholds used have been included, and the techniques justified in more
> detail.

3. P6 l30-31, in the phone shop example, the authors mentioned that it is
   possible to identify background noise based on a sharp rise of the request
   number; however, this will also eliminate the regular pedestrian flow in that
   shop and cause additional inaccuracy. Please clarify.

> Thank you. Clarified in the revised version. It is know with preliminary
> investigation that the received signal strength indicator decreases with
> increasing distance between the mobile device and the base station. This
> change is also found to be exponential as opposed to linear. With this context
> When we install a sensor to have a clear view of the pedestrian footpath in
> front of it, we make sure that all the devices beyond a certain distance will
> report similarly low signal strength. This threhold is idenfied using the
> kmeans algorithm to find the natural break in the data. We have found that
> this successfully isolates a farther stationary source of large number of
> probe requests such as a phone shop or a restaurant next door, Bustops etc.
> This doesn't work when the source of error is within the range of the
> footfall. As shown in our examples 2,4. Here our methodology fails to provide
> sufficient accuracy.

4. P7 l21, how often does the recycling of sequence number occur, and how will
   this impact the reliability of your analysis? Please provide more details.

> The recycling of the sequence number occurs every 4096 counts. From our
> experiments we observed that a device staying next to device actively
> searching recycles the sequence number 1-2 times in a 30 mins interval. On
> average this is expected to be much less than this. when the device doesn't
> probe as frequently. We have included more details regarding the cycling of
> sequence numbers and have mentioned method to overcomet this as mentioned by
> other reviewer. The change in implementing the method is estimated to be --%
> which we found to be not significant in our context.

5. P7 Section 3.3, please be more specific about the calibration process and the
   external source of information to be used here.

> This section has been expanded to include more details. The calibration
> process is adjusting the future counts based on a sample manual counts done at
> the location. The external source of information used is the manual counts
> which is collected outside the sensor

6. P7 l45, the analysis in this research is based on one single sensor for each
   location; however, in reality, it is very common to estimate travel flows
   based on multiple sensors (e.g., nearby stores in a shopping mall). How would
   your method address duplication in multiple sensors?

> This is a valid concern. Which we tried to clarify in the revised manuscript.
> Our motivation is not to attempt to track the movements of individuals,
> although we do attempt to accommodate measurement variations attributable to
> micro-site location of the sensors. In related work we have carried out
> controlled comparison of sensor measures obtained from adjacent store
> locations, and examined correspondence with retail unit function. This enables
> identification of the differences in dwell time between, for example, fast
> food outlets and clothes shops. This full analysis requires addition of store
> till data and the analysis is detailed. For this reason we include only a
> short reference to this extension of our analysis here. We also aim this
> method to count footfall accurately at locations rather than the flow of
> footfall between those locations. But as demonstarted in

> Lee J, Nemati S, Silva I, Edwards BA, Butler JP, Malhotra A. Transfer entropy
> estimation and directional coupling change detection in biomedical time
> series.  Biomed Eng Online. 2012;11:19. Published 2012 Apr 13.
> doi:10.1186/1475-925X-11-19

> and

> Vicente R, Wibral M, Lindner M, Pipa G. Transfer entropy--a model-free measure
> of effective connectivity for the neurosciences.  J Comput Neurosci.
> 2010;30(1):45-67.

> we can infer the relation or flow between the locations using other
> probabilistic approaches.

7. P10 l48, how was the adjustment factor determined?

> The adjustment factor is calculated as the ratio between the manual counts to
> the counts calculated from sensor for the sample period. This is used to
> adjust future counts at the same locations. We have included the clarification
> on the revised manuscript.

4 Reviewer 3
------------

4.1 Overall Summary
-------------------

1. This paper presents Wi-Fi based footfall counting methodology, which is
   timely and interesting topic. The method can be performed to estimate human
   activity such as pedestrian footfall from Wi-Fi probe requests.

> Yes, Thank you for your encouraging comment.

4.2 General Comments
--------------------

Some areas where I would like to see more detail:

1. It would have been interesting to understand the temporal characteristics for
   a week pedestrian footfall. How to consider this situation in your study? Why
   do you choose time range from 12:30 to 13:00 hrs ?

> Thank you. Our related work (cited below) examines daily, weekly and seasonal
> variation, and wil be the subject of future papers.

> Lugomer, K., & Longley, P. (2018). Towards a Comprehensive Temporal
> Classification of Footfall Patterns in the Cities of Great Britain (Short
> Paper). In LIPIcs-Leibniz International Proceedings in Informatics (Vol. 114).
> Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik.

> There was no reason for selecting the time in the pilot study. the location
> was chosen for it it known to have high number of randomised probe requests
> and has one of the highest footfall in London. For the main study the counting
> was done such that each location has atleast 2/3 counts in the same time
> interval.

2. Will you give more detailed information about the data and the parameters for
   various clustering algorithms? Including the number of records, the number of
   clusters. How about the performance for different clustering methods in your
   study？

> Thank you. This part has been expanded and clarified in the revised
> manuscript.  The performance of the algorithms used for one dimensional
> clustering were evaluated on the basis of the time taken for processing and
> the resulting mean absolute percentage error. We found that the k-means gives
> us the best data for the quickest processing time. The graph based clustering
> was done using two parameters - sequence threshold of 60 and time threshold of
> 16 seconds. These values were arrived arrived at using trial and error by
> using a know device as the control mechanism. This has been illustrated
> further in the revised manuscript.

4.3 Corrections and Clarifications
----------------------------------

More specific comments can be found in the following.

1. Minor formatting issues and typos needs to be fixed.  Figure 2, no value and
   units in the x-axis and y-axis.  Table 1, and Figure 4 are in wrong location.

> Thank you. This has been corrected in the revised version of the manuscript.
> Figure 2 is just a schematic showing the idea of cut off and clustering hence
> there are now values on the axes. We have included the units in the
> description

2. Related work: How about WiFi-based real-time data analytics?

> We have included reference to such work in our extended discussion of the
> relevance of the work to the smart cities discourse.

3. Methodology: Will you draw a diagram of the Methodology?

> We have resequenced some of the material in response to Referee 2 and,
> cognisant of over-all length requirements, think that this (rather than a
> diagram) addresses your concerns.

4. Performance Analysis: How about the performance for collecting data using
   PostgreSQL? Did you compare Apache Flink and Spark Streaming for real-time
   data collecting?

> This is an interesting point in the context of big data and tools for tackling
> big data. Though the data set here is large enough to make desktop computing
> unfeasible, it is neither big nor fast enough for a large scale big data
> system. An opensource and free toolkit comprising of text processing tools and
> relational database management system was enough to handle the compexity, size
> and speed of the data.  The toolkit employed here involves PostgreSQL, Bash,
> Unix tools for text processing (sed, awk, sort, join), Node.js and R.  The
> same toolkit proved to be useful in the larger project (smart street sensor)
> providing near real-time delivery of data.

5. The reference list needs to be written in a standard format. Some reference
   papers are lacking specific pages .Please use a standard way of reference
   list.

> Thank you. This has been corrected in the revised version of the paper.