### 2018-12-03 paper/sections/introduction.tex
intro and conclusions corrected

New and developing ‘smart’ technologies today provide the infrastructure over
which movements and interactions of people can be measured and monitored in the
‘sentient city’ \citep{amin2017seeing}. This is making it possible to
reinvigorate conceptualisations of city as the locus of human activities
supplementing night time geographies of residence \citep{martin2015developing}
with geographies of shopping behaviour \citep{lloyd2018detecting}, workzone
geographies \citep{singleton2018london} and studies of movement trajectories
\citep{campbell2008transforming}. This is rendering activity-based conceptions
of human behaviour central to the analysis of hardship and opportunity in, and
around, the smart city \citep{venerandi2015measuring}.

Sentient technologies include mobile phone networks, which can triangulate user
locations relative to networks of masts, the use of GPS to locate users of
social media services, and Wi-Fi access points providing internet access.  These
technologies offer differing levels of spatial precision, where mobile telephony
and Wi-Fi are generally less reliable and offer lower precision than GPS to the
end users, whilst simultaneously being more advantageous for broader mobility
studies \citep{pinelli2015comparing}. There has been considerable research into
the utility of these technologies to understand patterns of movement in cities
in near real time \citep{candia2008uncovering, gonzalez2008understanding,
calabrese2013understanding}.  Most of this research has focused upon technical
specification of accuracy or precision \citep{song2010limits,lane2010survey},
with somewhat less attention devoted to the ways that the characteristics of the
technologies and of their human users, conspire to create possible bias in
representing usage patterns across the entire smart city. Analysis of mobile
phone data, usually derived from industry players that have significant market
share and user bases representative of local populations, may also exclude
groups such as tourists from distant origins or subscribers to third party
services that share distinctive characteristics \citep{di2016mind}.

These examples illustrate the issues that underpin the assembly and analysis of
consumer data. Consumer data can be considered as a distinctive class of Big
Data which arise from the interactions between humans and customer-facing
organisations such as retailers, domestic energy suppliers, transport providers
and suppliers of social media and communications \citep{cdrc2018consumer}.
Consumer data account for an ever-increasing real share of all of the data that
are collected about citizens, but a fundamental characteristic of consumer-led
markets is that no single provider has a monopoly in market provision; therefore
issues of market share and segment generate bias in the analysis.  The source
and operation of this bias is unknown in the absence of extensive and context
sensitive attempts to triangulate consumer data with data of known provenance
relating to clearly defined populations \citep{lansley2016deriving}.  In similar
ways to other classes of Big Data, consumer data are best thought of as digital
‘exhaust’, or a by-product created by, or harvested from, consumer transactions.

In this paper, through a set of experiments, we evaluate the value of data
collected a network of 800 devices \citep{sss2016} installed across Great
Britain in order to characterize the footfall patterns of a scientifically
balanced sample of retail centres.  These devices are located in shop windows,
and record the probes emitted by mobile phones and other Wi-Fi enabled devices.
The data collected from these devices are deemed to be consumer data because
devices carried by consumers routinely probe for Wi-Fi connection which is a
consumer service.  Monitoring the probes from such devices provides an
indication of the presence of their users, regardless of whether or not internet
connectivity is established.  Our core motivation is to appraise the usefulness
of Wi-Fi probe requests harvested from our network of sensors as a method of
indicating levels of pedestrian activity.  More broadly still, in our future
research we intend to classify the nationwide network of footfall profiles as
part of a programme of research to understand the form and function of retail
areas at a time of far-reaching structural change for the retail industry.

To this end, it was important to first undertake a thorough conceptual and
technical appraisal of our consumer data source.  In technical terms, screening
the information present in the `probe requests' and clustering them based on
their characteristics was essential in order to remove those emitted by devices
which do not indicate pedestrian activity, such as network enabled printers and
other fixed devices.  Related to this, a method to fingerprint Wi-Fi probes was
necessary in order to remove probes from individuals’ devices that in conceptual
terms should not be considered part of footfall; for instance when , an employee
is seated in an office within range of the sensor device.  A calibration of
sensor measurement was also essential on two grounds: first individuals may
carry multiple devices, or no device at all; and second, the positioning and
orientation of the sensor in the retail unit may lead to systematic over- or
under-enumeration.  These sources of bias in measurement must be accommodated by
manual recording of footfall at each location and the generalization of these
sample survey results to all locations and time periods.  As we describe in
detail below, manual validation of the data needed to be undertaken in parallel
with the technical profiling of the mix of consumer mobile devices that probed
our sensors, since the effectiveness of data cleaning procedures discussed in
this paper differ between individual locations and configurations.