### 2018-12-07 paper/sections/previous_work.tex

In the past decade, Wi-Fi has emerged as one of the most commonly used
technologies in providing high speed internet access to mobile devices such as
smartphones, tablets and laptops in public and private spaces. This has
resulted in multiple Wi-Fi networks being available at almost every location in
dense urban environments. Traversing through this overlapping mesh of Wi-Fi
networks, modern mobile devices with Wi-Fi network interfaces regularly
broadcast a special type of signal known as `Probe Requests' in order to
discover the Wi-Fi networks available to them. This helps these devices to
connect and switch between the Wi-Fi networks seamlessly.

Probe requests are low level signals standardised by IEEE 802.11 specification
\citep{ieee2016} for service discovery, and are implemented in any Wi-Fi
capable device irrespective of the manufacturer or the model. This ubiquity and
standardisation makes them an excellent source of open, passive, continuous,
and wireless data generated by Wi-Fi capable devices present at any given time
and location. Considering the unprecendented levels of mobile device ownership
in recent years, we can, in turn use this data to understand the population
distribution in highly dynamic urban environments with high spatial and
temporal granularity \citep{freud2015, konto2017}. While a Wi-Fi based method to
collect data offers us various advantages such as, easy scalability and
efficiency in terms of cost and time, it also introduces few systematic biases
and uncertainities in the collected data along with the serious risk of
infringing on the privacy of the mobile users. In this paper, using a set of
probe requests and manual counts collected at various high street locations
across London, we demonstrate that pedestrian footfall at these locations can
be estimated with considerable precision and accuracy while protecting the
privacy of the pedestrians.

Unlike GPS, the location of the Wi-Fi enabled mobile device cannot be directly
inferred from Wi-Fi, however there are reliable methods to triangulate the
location of mobile devices from the locations of known access points (AP) and
the signal strength reported by them \citep{he2003range, moore2004robust,
lamarca2005place}. This can overcome the usual shortcoming of GPS, which
struggles for precision and accuracy in indoor and densely built environments
\citep{zarim2006, kawaguchi2009wifi, xi2010locating}. Utilising this, we can
easily and quickly estimate trajectories of the mobile devices
\citep{Sorensenlocation, musa2012tracking} which can be used similary to the
GPS trajectories to understand individual travel patterns \citep{reki2007,
Sap2015}, crowd behaviour \citep{abedi2013bluetooth, mowafi2013tracking},
vehicular \citep{lu2010vehicle} and pedestrian movement
\citep{xu2013pedestrian, fukuzaki2014pedestrian, wang2016gait}. Such data can
also be used in transportation planning and management to estimate travel time
\citep{musa2011wiflow} and real time traffic monitoring
\citep{abbott2013empirical}. Using techniques demonstrated by
\cite{franklin2006passive} and \cite{pang2007802}, along with information
present in the probe requests, one can even model interactions between the
users \citep{cheng2012inferring, barbera2013signals, cunche2014know,
cunche2014linking} such as predicting which of them are most likely to meet
again \citep{cunche2012know}. Using the semantic information present in these
probe requests it even is possible to understand the nature of population at a
large scale \citep{di2016mind}.

Although extensive research has been carried out on this subject with feasible
and favorable results, in recent years, one of the major challenges faced in
such attempts has been the increasing attempt by mobile phone manufacturers to
protect their users’ privacy by anonymising the globally identifiable portion
of the probe requests \citep{green2008}. Various methods have been devised to
overcome this anonymisation process such as estimating the device model
information from a known dataset of manufacturers and device behaviours
\citep{martin2016decomposition}; Scrambler attack using a small part of the
physical layer specification for Wi-Fi \citep{vo2016, bloessl2015scrambler};
and timing attack where the packet sequence information along with information
elements present in the probe request frame is used \citep{matte2016,
cheng2016can}. A combination of these methodologies has been proven to produce
de-anonymised globally unique device information \citep{vanhoef2016,
martin2017}. These approaches usually result in serious risk of breach of
privacy of the users of the mobile devices by revealing their
identifiable personal information.

There is a clear gap in the research for exploring methodologies for estimating
the number of unique mobile devices from a set of anonymised probe requests,
without the need to reveal their original device information. Such a technique
has various applications such as uncovering the urban wireless landscape
\citep{rose2010mapping}, revealing human activity at large scales
\citep{qin2013discovering}, estimating pedestrian numbers in crowds
\citep{schauer2014estimating, fukuzaki2015statistical}, and even counting
people in hyper local scales such as queues \citep{wang2013measuring}. With
enough infrastructure to collect such information we can even aim to generate a
real-time census of the city \citep{konto2017}. With this background, we set
out to devise and implement a methodology to reliably estimate human activity
such as pedestrian footfall from Wi-Fi probe requests without risking a breach
of privacy of the users involved.