### 2019-01-21 review/review_02/response.md
almost done

## Reviewer 1
--------------------------------------------------------------------------------

I appreciate the authors’ effort to address my comments in the revised draft.
The manuscript has been greatly improved. I have a few minor comments:

 1. p9 l36-40: When determining the threshold for high/low signal strength, did
    you just set the number of clusters equals to 2 for kmeans and hierarchical
    clustering? If so, you should explicitly say so in the explanation.

> Thanks for the feedback. We have included this explicitly in the revised
> manuscript.

 2. Please enlarge and recreate the sub-figures in Figure 3. This figure is
    difficult to read in its current format.

> We agree. A more legible version of Figure 3 has been included in the revised
> manuscript.

 3. I would suggest adding the following paper to the literature review:
    Torrens, P. M. (2008) Wi-Fi Geographies. Annals of the Association of
    American Geographers, 98(1), 59-84.

> Thanks for pointing out this important literature that we missed to cite.  We
> have included this reference in the literature review.

## Reviewer 2
--------------------------------------------------------------------------------

The revised version of this paper is of better quality and the authors have
taken into account reviewers' comments.

 1. However, Figure 3 needs a slight improvement. In the final version, all the
    figures must be of good quality, with similar formatting (size).

> We agree. A more legible version of Figure 3 has been included in the revised
> manuscript.

 2. Please check the format of the references.

> Thanks for pointing this out. We have checked and fixed the format of the
> references

More specific comments can be found in the following,

 3. Page 8, Could you give more description about your real-time experimental
    data? (Like data size, the total number of records)

> Thank you for the feedback. Similar to the Oxford street case study, we have
> included more descriptive statistics on the data collected under the case
> study in Section 5 - Table 2.

 4. Page 10, “we observed that the best thresholds for time and sequence
    numbers in the clustering algorithm is around 16 and 60 respectively”,
    Could you give more detail information?

> Thank you for the feedback. These thresholds are arrived at through a trial
> and error. We have further clarified the process along with a more legible
> illustration which is shown in Figure 3.

 5. Page 14. “There are obvious extensions to understanding the ebbs and flows
    of activities in the 24-hour smart city.” Could you give some suggestions
    to use your method in these extensions?

> Thanks for the input. We have included few examples and cited recent
> research demonstrating them.