### 2019-01-21 paper/sections/main_study.tex

The methodology set out above was implemented in five different Central London
locations at different times. Sensors were installed and data collected for
extended periods of time. We also carried out manual counting at these locations
across different times of the day. We then applied the methodologies discussed
earlier to arrive at estimated pedestrian footfall and compared them with the
corresponding manual counts.  We finally evaluated the effectiveness of the
processes with the Mean Absolute Percentage Error (MAPE) at the locations and
report our findings below.

\begin{table}
    \tbl{Locations where sensors were installes, volume and speed of probe requests collected by
    the sendor and total pedestrians manually counted. The data occupies around 1.8 GB on disk 
    when encoded in text format.}
	{\begin{tabular}{clllccc} 
		\toprule
         ID & Location & Type & Installation notes & Probe Requests & Footfall\\
         & & & & x10\textsuperscript{6} (per min) & No. (per min)\\
		 \midrule
         1 & Camden High Street & Phone Shop & Bus stop in front & 9.9 (297) & 3683 (33)\\
         2 & Central St.Giles & Restaurant & Seating area on both sides & 3.9 (169) & 0346 (05)\\
         3 & Holborn Station & Info. Kiosk & Overlooks station entrance & 4.3 (303) & 2956 (46)\\
         4 & Brunswick Center & Fast Food & Has seating area on one side & 3.4 (210) & 0960 (12)\\
         5 & The Strand & Tea Shop & Has phone shop next door & 8.4 (382) & 1969 (21)\\
		 \bottomrule
	\end{tabular}}
	\label{locations-table}
\end{table}

\begin{figure}
	\begin{center}
		\includegraphics [width=0.90\linewidth] {images/main_schedule.jpeg}
		\caption{Data collection schedule showing the days when sensors were active at their corresponding locations. The red squares show that manual counting of pedestrians was also done on that day.}
		\label{main_schedule}
	\end{center}
\end{figure}

The locations at which the data were collected are shown in Table
\ref{locations-table}. The locations were chosen for their diverse site
conditions and unique sources of noise around the potential location of the
sensors. The position of the sensor at these locations with respect to the
context is shown the Figure \ref{main_signal_strength}. We can see that
Location 5 is the `cleanest' with one clear stationary source of noise (phone
shop) while location 2 is the most complex due to the proximity of seating
areas to the sensor.  The sensors were operational through out February and
March, while manual counts were conducted in these locations in half hour
sessions on at least two different days. For the purposes of comparing with
ground truth, we considered the data from sensors which correspond to the 12
sets of available manual counts. The schedule of data collection is shown in
Figure \ref{main_schedule}.

\begin{figure}
	\centering
	\subfigure[Installation configuration of sensors at the survey locations (not to scale)]{
		\resizebox*{0.46\linewidth}{!}{\includegraphics[trim=20 6 20 6,clip]{images/main_configs.jpeg}}}\hspace{20pt}
	\subfigure[Density distribution of signal strength reported in collected probe requests (lower values show higher signal strength)]{
		\resizebox*{0.46\linewidth}{!}{\includegraphics[trim=20 4 25 6,clip]{images/main_signals.jpeg}}}
	\caption{Distribution of signal strengths across locations} \label{main_signal_strength}
\end{figure}

We begin by looking at the distribution of the signal strength reported by the
probe requests across the locations. From the density plot shown in Figure
\ref{main_signal_strength}, we can observe that there is clear relation between
the distribution of the signal strength and the distance and complexity of the
source of noise. We can see that while location 5 shows clean difference
between low and high signal strengths, location 2 is almost normally
distributed.  Intuitively we expected that location 2 and 4 would be harder to
classify than locations 1, 3 and 5.  We ran the k-means clustering algorithm
and filtered out the probe requests which were randomised and had signal
strengths less than the second break (threshold).  It is important to note that
we were dealing with relative thresholds of signal strengths which can vary
with location and time of the analysis.  We then aggregated the probe requests
by counting the number of Unique MAC addresses present in every minute. We also
removed devices that dwelled around the sensor by removing the MAC
addresses which reappeared within the previous hour.

The results of the first stage of the filtering process along with the
thresholds are shown in Table \ref{errors_table}.  Confirming our intuition, we
see that the location 2 has the most MAPE followed by location 4, while the rest
of them have highly reduced MAPE.  It is significant that this method alone
reduces our margin of error by 50 - 100\% from the raw counts without any
cleaning.  This makes the signal strength filtering a quick and ideal method for
practical applications, one which doesn't require absolute numbers such as
creating large aggregated indexes to show long-term trends.  We also found that
the success of the signal strength filtering can be improved significantly by
installing sensors so that the pedestrians and source of noise are at different
distances from the sensor. This ensures that the distribution of signal
strengths within the field of measurement is distinct from that of the
surroundings.

We then ran the sequence numbers based clustering process on the rest of the
probe requests to reduce the MAPE by almost 50 - 100\% on all the sensors except
for location 3.  Location 3 is an outlier among all the other sensors since it
is the only one with large amount of pedestrians very close to the sensor.  This
may be the reason behind the over filtering observed in the previous process.
We finally ran the calibration process where we calculated the adjustment
factors from the ratio between the manual counts to the sensor based counts for
the sample period as shown in Table \ref{errors_table}.  We used them to adjust
the counts to achieve a MAPE ranging from 10 - 50\%. We observed that the
sensors with people moving right next to them tend to under-count with our
methodology, while sensors with seating next to them tend to over-count
significantly. However, using the filtering process, we can reduce the error to
almost 10\% closer to that of the ground truth.

\begin{table}
	\tbl{Results of footfall estimation at each location as Mean Absolute Percentage Error (MAPE) after each step of the filtering process}
	{\begin{tabular}{ccccccc} 
		\toprule
			Sensor & Signal strength & Adjustment & MAPE & MAPE after & MAPE after & MAPE of\\
			& threshold & factor & without any & filtering signal & filtering sequence & final adjusted\\
			& (-dBm) & &  cleaning (\%) & strength (\%) & numbers (\%) & counts (\%)\\
		 \midrule
			1 & -70 & 1.25 & 259 &  22 & -13 &  9 \\
			2 & -74 & 0.51 & 928 & 396 & 206 & 55 \\
			3 & -72 & 1.60 &  87 & -19 & -31 & 10 \\
			4 & -70 & 0.88 & 498 & 142 &  52 & 33 \\
			5 & -72 & 0.80 & 473 &  84 &  38 & 11 \\
		 \bottomrule
	\end{tabular}}
	\label{errors_table}
\end{table}

\begin{figure}
	\begin{center}
		\includegraphics [width=\linewidth,trim=6 6 6 6,clip] {images/main_comparison.jpeg}
		\caption{Comparison of the filtering process with the ground truth in all the locations.}
		\label{main_comparison}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics [width=\linewidth] {images/main_footfall.jpeg}
		\caption{A week of pedestrian footfall at the Strand, London collected by the methodology. The counts are aggregated for 5 minute intervals.}
		\label{main_study_counts}
	\end{center}
\end{figure}