### 2019-01-22 stages/second_submission/sections/abstract.tex

The accurate measurement of human activity with high spatial and temporal
granularity is crucial for understanding the structure and function of the built
environment.  With increasing mobile ownership, the Wi-Fi `probe requests'
generated by mobile devices can act as a cheap, scalable and real-time source of
data for establishing such measures.  The two major challenges we face in using
these probe requests for estimating human activity are: filtering the noise
generated by the uncertain field of measurement, and clustering anonymised probe
requests generated by the same devices together without compromising the privacy
of the users.  In this paper we demonstrate that we can overcome these
challenges by using class intervals and a novel graph based technique for
filtering and clustering the probe requests which in turn, enables us to
reliably measure real-time pedestrian footfall at retail high streets.