### 2019-01-22 stages/second_submission/sections/conclusion.tex

Sentient technologies make measurement of the human activities that are the life
blood of the smart city possible. Yet the data that they harvest are frequently
relevant only to the sub-groups within society that avail themselves of
particular goods and services – such as social media applications, transport
modes or retail offers. In each of these cases, it is necessary to remember that
the resulting data are by-products of consumer transactions, and will as a
consequence, only pertain to users of the relevant goods or services. If the
smart city is to be socially inclusive, it therefore follows that sentient data
must represent entire populations, whether by design or by triangulation with
external, population wide, sources. This is a non-trivial task, since the ebbs
and flows of smart device-enabled citizens rarely pertain to any clearly defined
population in either administrative or functional terms
\citep{massam1975location}.

Our objective here has been to collect, rather than re-use, data on smart city
functioning, by recording Wi-Fi probes and ultimately reconciling them with
manual counts in order to infer ambient populations. The internal validation
methodology set out in the technical sections of this paper, allied to external
validation from pedestrian counts, renders the method inclusive and robust when
recording activity levels in retail centres in real time. We have described the
collection and processing of a novel consumer Big Dataset that enables valid
measures of levels of footfall activity which has been scaled across a wide
network of sensors \citep{cdrc2018consumer}.  In both conceptual and technical
terms, it illustrates the ways in which passively collected consumer data can be
‘hardened’ to render them robust and reliable by using related procedures of
internal and external validation.

Internal validation addresses the issues of screening out device probes that do
not indicate footfall, and the further screening of device probes to
‘fingerprint’ the effects of MAC randomization. It is important to note that the
filtering process work based solely on the information present in the probe
requests and their temporal distribution. This ensures that although the mobile
devices were uniquely identified, there was no further personal data generated
by linking the probe requests to the users of the mobile devices. This method
essentially gave us a way to estimate the footfall in real-time without
identifying or tracking the mobile devices themselves. External validation then
entailed reconciling adjusted counts with the footfall observed at sample
locations. This procedure makes it possible to generalise from locations at
which manual footfall surveys are conducted to all others in the system, and to
develop a classification of device locations that are more or less susceptible
to noise generation.

This Wi-Fi based footfall counting methodology offers a large number of
applications and benefits for real time spatial analysis. Since Wi-Fi based
sensors are inexpensive and the data model is scalable, it is possible to use
this methodology for a large network of sensors to gather granular data on
pedestrian footfall. A snapshot showing a week’s worth of precise footfall in
area around Charring cross, London is shown in Figure \ref{main_study_counts} in
order to demonstrate the potential for such a dataset. Projects such as
SmartStreetSensors \citep{cdrc2018consumer}, may utilise this methodology to
overcome the challenges introduced by the implementation of MAC address
randomisation.

The vicissitudes of MAC randomisation, and the provisions of privacy legislation
such as EU General Data Protection Regulations mitigate against tracking
individuals across the smart city using this approach.  This can be modelled
using agent-based methods \citep{heppenstall2011agent}, however. In our own
research we have also begun to link store time-lagged till receipts to footfall,
and have used such data to better understand the dwell times that characterise
such different retail uses as stores with window displays and fast food
restaurants.  Such analysis not only provides a more nuanced picture of movement
through retail areas, but also enables valorisation of micro sites within retail
centres. In the UK, for example, this is of immediate practical importance in
evaluating business rates on properties, and has still wider implications for
the setting of retail unit rental values. There are obvious extensions to
understanding the ebbs and flows of activities in the 24-hour smart city.

More broadly still, extensions to this strand of smart city research are likely
to seek to differentiate the quality of different elements within footfall
according to mission e.g. travel to adjacent workplace zones, leisure, etc., and
personal characteristics such as spending power. In this respect, future
research may not only simulate linkage of harmonised footfall counts between
sensor locations, but also link these in turn to disaggregate origin-destination
matrices for bikeshare and other public transport modes.  Our own investigations
will consider these and other challenges to understanding the functioning of the
sentient city.